# Trace Element Analysis

```{r}
#| include: false

set.seed(79)

ref <- data.frame(
  element = c("Mg","Ca","V","Cr","Mn","Fe","Ni","Cu","Zn","Sr","Ba",
              "La","Ce","Pr","Nd","Sm","Eu","Gd","Tb","Dy","Ho","Er",
              "Tm","Yb","Lu","Pb","Th","U"),
  mean = c(3075, 391417, 0.028, 0.8, 1.95, 15, 2.2, 3.6, 215, 168, 4.6,
           0.007, 0.010, 0.0014, 0.006, 0.004, NA, NA, 0.0005, 0.0015,
           0.0005, 0.0014, 0.0003, 0.0026, 0.0005, 6.55, 0.0031, 0.003),
  sd   = c(623, 18230, 0.014, 1.83, 1.72, 19, 4.9, 7.4, 77, 178, 4.4,
           0.012, 0.015, 0.002, 0.008, 0.005, NA, NA, 0.0003, 0.0015,
           0.0004, 0.0013, 0.0002, 0.0031, 0.0004, 8.6, 0.006, 0.005),
  min  = c(1931, 350011, 0, 0, 0.51, 2.23, 0.11, 0.01, 80.4, 58, 0.52,
           0, 0, 0, 0, 0, NA, NA, 0, 0, 0, 0, 0, 0, 0, 0.04, 0, 0),
  max  = c(5225, 425928, 0.079, 10.3, 12, 104, 36.5, 32.9, 474, 1344, 31.3,
           0.079, 0.087, 0.012, 0.042, 0.020, NA, NA, 0.0017, 0.0056,
           0.0012, 0.0057, 0.0007, 0.0148, 0.0017, 45.5, 0.035, 0.036),
  mtc  = c(6472, 462387, 0.11, 13.9, 15.4, 143, 46, 48, 629, 1700, 40,
           0.1, 0.12, 0.015, 0.058, 0.030, NA, NA, 0.002, 0.009, 0.002,
           0.008, 0.001, 0.021, 0.003, 63, 0.05, 0.05)
)

simulate_element <- function(m, s, lo, hi, n = 10) {
  if (is.na(m)) return(rep(NA_real_, n)) 
  sigma2 <- log(1 + s^2 / m^2)
  mu     <- log(m) - sigma2 / 2
  out <- numeric(0)
  while (length(out) < n) {
    x <- rlnorm(n, mu, sqrt(sigma2))
    out <- c(out, x[x >= lo & x <= hi])
  }
  signif(out[1:n], 3)
}

enamel <- as.data.frame(
  mapply(simulate_element, ref$mean, ref$sd, ref$min, ref$max,
         SIMPLIFY = FALSE),
  col.names = ref$element
)
enamel <- cbind(sample = sprintf("ME%02d", 1:10), enamel)
```

## Checking for Diagenesis {.unnumbered}

### Maximum Threshold Concentration (MTC) {.unnumbered}

Source: Kamenov et al., 2018

```{r}
#| message: false

library(gt)
library(dplyr)

Elements <- c("Mg", "Ca", "V", "Cr", "Mn", "Fe", "Ni", "Cu", "Zn", "Sr",
              "Ba", "La", "Ce", "Pr", "Nd", "Sm", "Eu","Gd" , "Tb", "Dy",
              "Ho", "Er", "Tm", "Yb", "Lu", "Pb", "Th", "U")

MTC <- c(6472, 462387, 0.11, 13.9, 15.4, 143, 46, 48, 629, 1700,
         40, 0.1, 0.12, 0.015, 0.058, 0.030, NA, NA, 0.002, 0.009,
         0.002, 0.008, 0.001, 0.021, 0.003, 63, 0.05, 0.05)

trace <- data.frame(Elements, MTC)
```

### Normalizing your data (divide by the MTC) {.unnumbered}

1.  Transform data into a wide format

    Depending on the format of your data, transform the dataframe so that the rows represent the samples/individuals and the columns represent the trace elements.

2.  Normalize each column by dividing by the MTC

```{r}
mtc <- trace$MTC[match(names(enamel)[-1], trace$Elements)] # ensures the elements are in the same order
mtc <- trace$MTC[match(names(enamel)[-1], trace$Elements)] 
# ensures the elements are in the same order
# if you have additional variables, exclude those as well in place of [-1] which removes the first column of sample names

mtc.ratio <- cbind(sample=enamel$sample, sweep(enamel[,-1], 2, mtc, "/")) 
# divides the sample values by the matched MTC values
# sweep(x, Margin, Stats, Fun)
```

3.  Create a subset of Ratios

```{r}
library(tidyr)

elements <- c("V", "Mn", "Fe", "La", "Ce", "Nd", "Dy", "Yb", "Th", "U")

mtc.long <- mtc.ratio %>%
  select(sample, all_of(elements)) %>% 
  # pulls the sample numbers and the values for the subset of elements
  pivot_longer(-sample, names_to = "element", values_to = "ratio") 
  # transforms the table into a stacked format
```

4.  Plotting the Normalized Concentrations

```{r}
library(ggplot2)

ggplot(mtc.long, aes(x=sample, y=ratio, group=element))+
  geom_point(aes(color=element))+
  geom_line(aes(color=element))+
  scale_y_log10() +
  geom_hline(yintercept = 10^0, color="red")+
  labs(x="Samples", y="C/MTC")
```

5.  Interpret the Data

Dividing the sample concentrations by the maximum threshold concentrations obtained from modern individuals acts as a diagenesis check. If any value exceeds the 10\^0 line, there is potential diagenic alteration to the sample.

## Dietary analysis {.unnumbered}

### Biopurification of Calcium {.unnumbered}

Ba/Ca vs. Sr/Ca

- Biopurification = preferentially assimilating essential calcium to non-essential strontium or barium in bones and teeth \* Burton et al. 1999
- Mammals only absorb 20-30% of dietary strontium \* Losch et al. 2014
- Barium is not soluble in water, sp Ba/Ca ratio is lower in marine foods \* Pankowska et al. 2016
- Sr/Ca and Ba/Ca ratios decrease in terrestrial trophic chains \*Balter 2004

d13C vs. Ba/Sr

```{r}
# 
```

Zn/Ca

```{r}
# Proportion of shellfish in diet
```
