# Types of Visualizations {.unnumbered}

## Distribution {.unnumbered}

```{r}
#| include: false
library(ggplot2)
data <- data.frame(value=rnorm(100, mean=15, sd=3))
data$group <- as.factor(sample(1:3, size=100, replace=TRUE))
```

### Boxplot {.unnumbered}

```{r}
#| label: basic-boxplot
#| fig-cap: "Basic Boxplot"

ggplot(data, aes(y=value))+
  geom_boxplot()
```

```{r}
#| label: group-boxplot
#| fig-cap: "Boxplot by Group"

ggplot(data, aes(y=value, x=group))+
  geom_boxplot()
```

### Violin {.unnumbered}

```{r}
#| label: violin
#| fig-cap: "Violin Plot by Group"
ggplot(data, aes(x=group, y=value))+
  geom_violin()
```

### Histogram {.unnumbered}

```{r}
#| label: basic-histogram
#| fig-cap: "Basic Histogram"
#| message: false

ggplot(data, aes(x=value))+
  geom_histogram()
```

```{r}
#| label: bin-histogram
#| fig-cap: "Histogram with designated Bins"

ggplot(data, aes(x=value))+
  geom_histogram(binwidth=.1)
```

### Density {.unnumbered}

```{r}
#| label: density

ggplot(data, aes(x=value))+
  geom_density()
```

## Correlation {.unnumbered}

### Scatter {.unnumbered}

### Bubble {.unnumbered}

### Heatmap {.unnumbered}

### Correlogram {.unnumbered}

## Map {.unnumbered}

### Map {.unnumbered}

### Choropleth {.unnumbered}

### Bubble Map {.unnumbered}

## Flow {.unnumbered}

### Chord Diagram {.unnumbered}

### Network {.unnumbered}

### Sankey {.unnumbered}

### Dendrogram {.unnumbered}
